---
title: "Bacterial Endophyte Project - Multivariate Analyses"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
author: "Briana K. Whitaker"
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning=FALSE, message=FALSE)
```

* Run using `r version[['version.string']] `.

# Load packages
```{r, echo = FALSE, results='hide'}
### packages
x <- c("tidyverse", "dplyr", "ggplot2", "car", "lme4", "RColorBrewer", "egg")  #"nlme", 
lapply(x, require, character.only = TRUE)

### functions and read-in source code
source("./code/multiplot.function.R")

get.pot <- function(fname) strsplit(fname, "[.]")[[1]][1] 
get.plant <- function(fname) strsplit(fname, "[.]")[[1]][2] 
get.pot2 <- function(fname) strsplit(fname, "[ ]")[[1]][1] 
get.plant2 <- function(fname) strsplit(fname, "[ ]")[[1]][2] 

## leaf phys functions
#Murray FW (1967) On the computation of saturation vapor pressure. 
#    J. Appl. Meteorol. 6: 203-204.
# requires air Temp as degrees in Celsius
# requires RH as a % relative humidity of the air surrounding the leaf
airVPD <- function(Temp, RH) {    
   SVP <- 610.78*exp((Temp*17.2694)/(238.3+Temp))
   VPD <- SVP*(1-RH/100)
   VPD
}
# Similar inputs as airVPD, but additional parameter required of leaf temperature
#  difference from ambient (i.e., "deltaT")
leafVPD <- function(Temp, RH, deltaT) {
   airSVP <- 610.78*exp((Temp*17.2694)/(238.3+Temp))
   leafSVP <- 610.78*exp(((Temp+deltaT)*17.2694)/(238.3+(Temp+deltaT)))
   leafVPD <- leafSVP-(airSVP*(RH/100))
   leafVPD
}

#
CV <- function(x) {
    x1 <- na.omit(x)
    MEAN <- mean(x1)
    SD <- sd(x1)
    CV <- MEAN/SD
    CV
}
myMEAN <- function(x) {
    x1 <- na.omit(x)  #number of omitted values length(attr(test,"na.action"))
    MEAN <- mean(x1)
}
mySD <- function(x) {
    x1 <- na.omit(x)  #number of omitted values length(attr(test,"na.action"))
    SD <- sd(x1)
}

myREP <- function(x) {
    x1 <- na.omit(x)  #number of omitted values length(attr(test,"na.action"))
    REP <- length(x1) #number of values used in mean
}

# micro treatment colors
# display.brewer.pal(8, "Dark2") #assumes Phal is first in order
# brewer.pal(8, "Dark2")
micro_color_neg <- c("#666666", "#1B9E77", "#D95F02", "#7570B3", "#E7298A", 
                 "#66A61E", "#E6AB02", "#A6761D", "black") 

micro_color <- micro_color_neg[-1]

# set ggplot2 theme
theme_set(theme_bw(base_size=12)) 
theme_update(panel.grid.major=element_line(0), panel.grid.minor=element_line(0))

```


# Load data
```{r}
disease3 <- read.csv("./data/GRIP2_disease_2022-07-13.csv", row.names = 1,
                     stringsAsFactors = TRUE)
disease4 <- read.csv("./data/GRIP2_diseaseNoNeg_2022-07-13.csv", row.names = 1,
                     stringsAsFactors = TRUE)
pot     <- read.csv("./data/GRIP2_Pot_Means_2022-08-09.csv", row.names = 1,
                     stringsAsFactors = TRUE)
phys4    <- read.delim("./data/GRIP2_phys_2022-07-13.txt", sep = " ", row.names = 1,
                     stringsAsFactors = TRUE)
```

```{r, echo = FALSE, results='hide'}
disease3$plantRep <- as.factor(disease3$plantRep)
disease3$potID <- as.factor(disease3$potID)
disease3$block <- as.factor(disease3$block)
disease3$microTrt <- factor(disease3$microTrt, 
    levels = c("NEG", "A30", "Q16", "Q22", "W14", "Y108", "Y11", "Y2", "POS"))

disease4$plantRep <- as.factor(disease4$plantRep)
disease4$potID <- as.factor(disease4$potID)
disease4$block <- as.factor(disease4$block)
disease4$microTrt <- factor(disease4$microTrt, 
    levels = c(       "A30", "Q16", "Q22", "W14", "Y108", "Y11", "Y2", "POS"))


phys4$potID <- as.factor(phys4$potID)
phys4$block <- as.factor(phys4$block)
phys4$bench <- as.factor(phys4$bench)
phys4$ID <- as.factor(phys4$ID)
phys4$plant_rep <- as.factor(phys4$plant_rep)
phys4$bacteria <- factor(phys4$bacteria, 
    levels = c("NO", "A30", "Q16", "Q22", "W14", "Y108", "Y11", "Y2"))
levels(phys4$bacteria)[1] <- "NONE"

pot$block <- as.factor(pot$block)

# str(disease3)
# str(disease4)
# str(phys4)
# str(pot)

```



# Remove effects of environmental covariates from pot-level Phys. vars.
* Remove Negative control treatment from this analysis
```{r, echo = FALSE, results = 'hide'}
pot$microTrt <- factor(pot$microTrt, levels = c("NEG", "A30", "Q16", "Q22", "W14", "Y108", "Y11", "Y2","POS"))

pot2 <- pot[pot$microTrt != "NEG",]
pot2 <- droplevels(pot2)
```

* Use BIC/SBC criterion, k = log(n), where n=48 pots
* Evaluating for ~ >2 difference in BIC

```{r}
phitwo.p0a <- lm(Phi2means ~ Tempmeans + PARmeans + lVPDmeans, data = pot2)
phitwo.p0b <- lm(Phi2means ~ 1, data = pot2)
phitwo.back <- step(phitwo.p0a, direction="backward", k = log(48), 
                    scope = c(lower = ~ 1))
phitwo.fwd <- step(phitwo.p0b, direction="forward", k = log(48),
                    scope = c(upper = ~ Tempmeans + PARmeans + lVPDmeans))
# both agree, only Tempmeans
phitwo.p1 <- lm(Phi2means ~ Tempmeans, data = pot2)
```

```{r}
phino.p0a <- lm(PhiNOmeans ~ Tempmeans + PARmeans + lVPDmeans, data = pot2)
phino.p0b <- lm(PhiNOmeans ~ 1, data = pot2)
phino.back <- step(phino.p0a, direction="backward", k = log(48), 
                    scope = c(lower = ~ 1))
phino.fwd <- step(phino.p0b, direction="forward", k = log(48), 
                    scope = c(upper = ~ Tempmeans + PARmeans + lVPDmeans))
#do not agree, but either one of Tempmeans or lVPDmeans (but not both) has substantial improvements to BIC in either direction, while PAR either has no effect or makes BIC worse
phino.p1 <- lm(PhiNOmeans ~ Tempmeans, data = pot2)
```

```{r}
phinpq.p0a <- lm(PhiNPQmeans ~ Tempmeans + PARmeans + lVPDmeans, data = pot2)
phinpq.p0b <- lm(PhiNPQmeans ~ 1, data = pot2)
phinpq.back <- step(phinpq.p0a, direction="backward", k = log(48), 
                    scope = c(lower = ~ 1))
phinpq.fwd <- step(phinpq.p0b, direction="forward", k = log(48), 
                    scope = c(upper = ~ Tempmeans + PARmeans + lVPDmeans))
# both agree, no predictors
phinpq.p1 <- lm(PhiNPQmeans ~ 1, data = pot2)
```

```{r}
npqt.p0a <- lm(NPQtmeans ~ Tempmeans + PARmeans + lVPDmeans, data = pot2)
npqt.p0b <- lm(NPQtmeans ~ 1, data = pot2)
npqt.back <- step(npqt.p0a, direction="backward",
                    scope = c(lower = ~ 1))
npqt.fwd <- step(npqt.p0b, direction="forward",
                    scope = c(upper = ~ Tempmeans + PARmeans + lVPDmeans))
# agree, lVPD
npqt.p1 <- lm(NPQtmeans ~ lVPDmeans, data = pot2)
```

```{r}
vH.p0a <- lm(vH.means ~ Tempmeans + PARmeans + lVPDmeans, data = pot2)
vH.p0b <- lm(vH.means ~ 1, data = pot2)
vH.back <- step(vH.p0a, direction="backward", k = log(48), 
                    scope = c(lower = ~ 1))
vH.fwd <- step(vH.p0b, direction="forward", k = log(48), 
                    scope = c(upper = ~ Tempmeans + PARmeans + lVPDmeans))
# agree, PAR
vH.p1 <- lm(vH.means ~ PARmeans, data = pot2)
```

```{r}
gH.p0a <- lm(gH.means ~ Tempmeans + PARmeans + lVPDmeans, data = pot2)
gH.p0b <- lm(gH.means ~ 1, data = pot2)
gH.back <- step(gH.p0a, direction="backward", k = log(48), 
                    scope = c(lower = ~ 1))
gH.fwd <- step(gH.p0b, direction="forward", k = log(48), 
                    scope = c(upper = ~ Tempmeans + PARmeans + lVPDmeans))
# agree, lVPD
gH.p1 <- lm(gH.means ~ lVPDmeans, data = pot2)
```

```{r}
ecs.p0a <- lm(ECStmeans ~ Tempmeans + PARmeans + lVPDmeans, data = pot2)
ecs.p0b <- lm(ECStmeans ~ 1, data = pot2)
ecs.back <- step(ecs.p0a, direction="backward", k = log(48), 
                    scope = c(lower = ~ 1))
ecs.fwd <- step(ecs.p0b, direction="forward", k = log(48), 
                    scope = c(upper = ~ Tempmeans + PARmeans + lVPDmeans))
# agree, PAR + lVPD
ecs.p1 <- lm(ECStmeans ~ PARmeans + lVPDmeans, data = pot2)
```

```{r}
LEF.p0a <- lm(LEFmeans ~ Tempmeans + PARmeans + lVPDmeans, data = pot2)
LEF.p0b <- lm(LEFmeans ~ 1, data = pot2)
LEF.back <- step(LEF.p0a, direction="backward", k = log(48), 
                    scope = c(lower = ~ 1))
LEF.fwd <- step(LEF.p0b, direction="forward", k = log(48), 
                    scope = c(upper = ~ Tempmeans + PARmeans + lVPDmeans))
# agree, Temp+ PAR (PAR VERY important to keep) -- LEF probably too dependent on PAR
LEF.p1 <- lm(LEFmeans ~ Tempmeans + PARmeans, data = pot2)
```

```{r, fig.width = 8, fig.height = 5}
par(mfrow = c(2,3))

plot(phitwo.p1$residuals)
anova(phitwo.p1) # temp signif
summary(phitwo.p1)$adj.r.squared #<0.10
coef(phitwo.p1)

plot(phino.p1$residuals)
anova(phino.p1) #temp signif
summary(phino.p1)$adj.r.squared #<0.20
coef(phino.p1)

plot(phinpq.p1$residuals)
anova(phinpq.p1)
summary(phinpq.p1)$adj.r.squared
coef(phinpq.p1)

# plot(npqt.p1$residuals)
# anova(npqt.p1) # lVPD not signif
# summary(npqt.p1)$adj.r.squared
# coef(npqt.p1)

plot(vH.p1$residuals)
anova(vH.p1) # PAR signif
summary(vH.p1)$adj.r.squared # <0.15
coef(vH.p1)

plot(gH.p1$residuals)    #slight trend
anova(gH.p1) # lVPD signif
summary(gH.p1)$adj.r.squared #<0.10
coef(gH.p1)

plot(ecs.p1$residuals)    
anova(ecs.p1) # PAR signif, lVPD semi-signif
summary(ecs.p1)$adj.r.squared #~0.20
coef(ecs.p1)

# plot(LEF.p1$residuals)
# anova(LEF.p1) # PAR very signif, temp signif
# summary(LEF.p1)$adj.r.squared  #!
# coef(LEF.p1)

par(mfrow = c(1,1))
```

```{r}
# # extra plots of LEF for checking dependencies
# plot(LEF.p1$residuals ~ pot2$LEFmeans)
# plot(pot2$LEFmeans ~ pot2$PARmeans) # extremely strong relationship
# plot(pot2$LEFmeans ~ pot2$Tempmeans)
```


```{r}
pot2 <- data.frame(pot2, 
        Phi2resids = summary(phitwo.p1)$residuals,
        PhiNOresids = summary(phino.p1)$residuals,
        PhiNPQresids = summary(phinpq.p1)$residuals,
        #NPQtresids = summary(npqt.p1)$residuals,
        ECStresids = summary(ecs.p1)$residuals,
        vHresids = summary(vH.p1)$residuals,
        gHresids = summary(gH.p1)$residuals       #,
        #LEFresids = summary(LEF.p1)$residuals
        )
```

```{r}
# sanity check of resids to orig values
plot(pot2$Phi2means ~ pot2$Phi2resids)

```


# PCA of Physiology Variables (not morph)

```{r, echo = FALSE, results = 'hide'}
#make an id for rownames
pot2$id <- paste(pot2$hostVar,pot2$microTrt,pot2$block, sep="-")
rownames(pot2) <- pot2$id


pot2 %>% filter(hostVar == 'Alsen') -> potAlsen
pot2 %>% filter(hostVar == 'Norm') -> potNorm
```

### VIFs
```{r}
M <- lm(Phi2resids ~., data = as.data.frame(pot2[,c(46:51)]))
vif(M) #gHresids highest
M <- lm(Phi2resids ~., data = as.data.frame(pot2[,c(46:50)]))
vif(M) #all <2 after dropping gH+


#PhiNOresids, PhiNPQresids, ECStresids, vHresids, gHresids

# not inclouded in VIFs leaf_thicknessmeans, SPADmeans
```

```{r, echo = FALSE, results = 'hide'}
# start with a dataset of all phys + morph traits
potAlsen %>% dplyr::select(Phi2resids, PhiNOresids, PhiNPQresids,
                       ECStresids, vHresids) -> potA_traits     #, SPADmeans, leaf_thicknessmeans
potNorm %>% dplyr::select(Phi2resids, PhiNOresids, PhiNPQresids,
                       ECStresids, vHresids) -> potN_traits    #, SPADmeans, leaf_thicknessmeans 
rownames(potA_traits) <- potAlsen$id
rownames(potN_traits) <- potNorm$id

#check correlations 
cor(potA_traits)
cor(potN_traits)
```

## Run PCA of phys
* Standardize data
```{r, echo = FALSE, results = 'hide'}
#standardize data
potA_traits2 <- scale(potA_traits)
potN_traits2 <- scale(potN_traits)
#run PCA
PCAtraitsA <- prcomp(potA_traits2, scale=TRUE)
PCAtraitsN <- prcomp(potN_traits2, scale=TRUE)
#PCAtraits$rotation #matrix of variable loadings
eigtraitsA <- (PCAtraitsA$sdev)^2; eigtraitsA    #get eigenvalues
vartraitsA <- eigtraitsA*100/sum(eigtraitsA); vartraitsA   
cumvartraitsA <- cumsum(vartraitsA); cumvartraitsA
vectorstraitsA <- as.data.frame(PCAtraitsA$rotation)*exp(1)

eigtraitsN <- (PCAtraitsN$sdev)^2; eigtraitsN    #get eigenvalues
vartraitsN <- eigtraitsN*100/sum(eigtraitsN); vartraitsN   
cumvartraitsN <- cumsum(vartraitsN); cumvartraitsN
vectorstraitsN <- as.data.frame(PCAtraitsN$rotation)*exp(1)
```

```{r}
# variation explained by axes
vartraitsA
vartraitsN

# loadings
PCAtraitsA$rotation
PCAtraitsN$rotation

```

* broken stick significance plot

```{r, echo = FALSE, results = 'hide'}
par(mfrow=c(1,1))
plot(I(PCAtraitsA$sdev^2), xlab = 'PCA Axis', ylab = 'Eigenvalue', las=1, 
     cex.lab=1.5, pch=16, main = "Alsen")
abline(h = mean(PCAtraitsA$sdev^2), lty=2, lwd=2,col='blue')
b.stick <- vegan::bstick(length(PCAtraitsA$sdev), tot.var=sum(eigtraitsA))
lines(1:length(PCAtraitsA$sdev), b.stick, type="l", lty=4, lwd=2, col='red') 
legend("topright", legend=c("Avg. Eigenvalue", "Broken Stick"), lty=c(2,4), 
       bty="n", col=c("blue", "red"))

plot(I(PCAtraitsN$sdev^2), xlab = 'PCA Axis', ylab = 'Eigenvalue', las=1, 
     cex.lab=1.5, pch=16, main = "Norm")
abline(h = mean(PCAtraitsN$sdev^2), lty=2, lwd=2,col='blue')
b.stick <- vegan::bstick(length(PCAtraitsN$sdev), tot.var=sum(eigtraitsN))
lines(1:length(PCAtraitsN$sdev), b.stick, type="l", lty=4, lwd=2, col='red') 
legend("topright", legend=c("Avg. Eigenvalue", "Broken Stick"), lty=c(2,4), 
       bty="n", col=c("blue", "red"))

```

* 3 sig axes for Alsen & Norm (Avg. Eigenvalue vs. Broken Stick Method)

```{r, echo = FALSE, results = 'hide'}
#merge with PCA with main dataframe
PCAtraits2A <- data.frame(ID=rownames(PCAtraitsA$x),PCAtraitsA$x)
PCAtraits3A <- merge(PCAtraits2A, potAlsen, by.x=c("ID"), by.y=c("id"), 
      all.y=TRUE)
rownames(PCAtraits3A) <- PCAtraits3A$ID
percentagetraitsA <- paste("PC-", as.character(seq(1,5,1)), " (",    
                     paste(as.character(round(vartraitsA,2)), "%", ")", sep=""), sep = "")

PCAtraits2N <- data.frame(ID=rownames(PCAtraitsN$x),PCAtraitsN$x)
PCAtraits3N <- merge(PCAtraits2N, potNorm, by.x=c("ID"), by.y=c("id"), 
      all.y=TRUE)
rownames(PCAtraits3N) <- PCAtraits3N$ID
percentagetraitsN <- paste("PC-", as.character(seq(1,5,1)), " (", 
                     paste(as.character(round(vartraitsN,2)), "%", ")", sep=""), sep = "")

```

### Alsen Phys PCA
```{r, echo = FALSE, results = 'hide', fig.keep = 'none'}
traitsPlotA <- 
    ggplot(PCAtraits3A, aes(x=PC1, y=PC2)) +
        coord_fixed() +  #ratio=5
        geom_point(size=3.5, aes(color=microTrt)) +  #alsen squares, norm circles?
        xlab(percentagetraitsA[1]) + ylab(percentagetraitsA[2]) +
        scale_shape_discrete("Variety") +
        scale_color_manual("",  values = micro_color) +
        theme(legend.position = "bottom", 
              axis.title = element_text(size = 13), axis.text = element_text(size = 13),
              legend.text=element_text(size=10), legend.title=element_blank()) +
        guides(color='none') +
        annotate(geom = "text", x = -2.5, y = 2.5, label = "A)\nAlsen", size = 5, hjust = 0)

traitsPlotVecA <- traitsPlotA + 
    geom_segment(aes(x=0, y=0, xend=vectorstraitsA[1,1], yend=vectorstraitsA[1,2]), data=vectorstraitsA, 
        color="#444444", size=1.2, arrow = arrow(length = unit(0.03, "npc"))) +
        annotate("text", x = vectorstraitsA[1,1]-0.35, y = vectorstraitsA[1,2]+0.15, label = "Phi2") +
    geom_segment(aes(x=0, y=0, xend=vectorstraitsA[2,1], yend=vectorstraitsA[2,2]), data=vectorstraitsA, 
        color="#444444", size=1.2, arrow = arrow(length = unit(0.03, "npc"))) +
        annotate("text", x = vectorstraitsA[2,1]-0.22, y = vectorstraitsA[2,2]+0.25, label = "PhiNO") +
    geom_segment(aes(x=0, y=0, xend=vectorstraitsA[3,1], yend=vectorstraitsA[3,2]), data=vectorstraitsA, 
        color="#444444", size=1.2, arrow = arrow(length = unit(0.03, "npc"))) +
        annotate("text", x = vectorstraitsA[3,1]+0.65, y = vectorstraitsA[3,2]+0.03, label = "PhiNPQ") +
    geom_segment(aes(x=0, y=0, xend=vectorstraitsA[4,1], yend=vectorstraitsA[4,2]), data=vectorstraitsA, 
        color="#444444", size=1.2, arrow = arrow(length = unit(0.03, "npc"))) +
        annotate("text", x = vectorstraitsA[4,1]-0.3, y = vectorstraitsA[4,2]+0.20, label = "ECSt") +
    geom_segment(aes(x=0, y=0, xend=vectorstraitsA[5,1], yend=vectorstraitsA[5,2]), data=vectorstraitsA, 
        color="#444444", size=1.2, arrow = arrow(length = unit(0.03, "npc"))) +
        annotate("text", x = vectorstraitsA[5,1]-0.3, y = vectorstraitsA[5,2]+0.10, label = "vH+")
    # geom_segment(aes(x=0, y=0, xend=vectorstraitsA[1,1], yend=vectorstraitsA[1,2]), data=vectorstraitsA, 
    #     color="#444444", size=1.2, arrow = arrow(length = unit(0.03, "npc"))) +
    #     annotate("text", x = gHVecA$x2-0.3, y = gHVecA$y2-0.05, label = "gH+")

traitsPlotVecA

```

```{r, echo = FALSE, results = 'hide', fig.keep = 'none'}
traitsPlotA2 <- 
    ggplot(PCAtraits3A, aes(x=PC1, y=PC3)) +
        coord_fixed() +  #ratio=5
        geom_point(size=3.5, aes(color=microTrt)) +  #alsen squares, norm circles?
        xlab(percentagetraitsA[1]) + ylab(percentagetraitsA[3]) +
        scale_shape_discrete("Variety") +
        scale_color_manual("",  values = micro_color) +
        theme(legend.position = "bottom", 
              axis.title = element_text(size = 13), axis.text = element_text(size = 13),
              legend.text=element_text(size=10), legend.title=element_blank()) +
        guides(color='none') +
        annotate(geom = "text", x = -2.5, y = 1.35, label = "C)\nAlsen", size = 5, hjust = 0)

traitsPlotVecA2 <- traitsPlotA2 + 
    geom_segment(aes(x=0, y=0, xend=vectorstraitsA[1,1], yend=vectorstraitsA[1,3]), data=vectorstraitsA, 
        color="#444444", size=1.2, arrow = arrow(length = unit(0.03, "npc"))) +
        annotate("text", x = vectorstraitsA[1,1]-0.1, y = vectorstraitsA[1,3]-0.15, label = "Phi2") +
    geom_segment(aes(x=0, y=0, xend=vectorstraitsA[2,1], yend=vectorstraitsA[2,3]), data=vectorstraitsA, 
        color="#444444", size=1.2, arrow = arrow(length = unit(0.03, "npc"))) +
        annotate("text", x = vectorstraitsA[2,1]+0.60, y = vectorstraitsA[2,3]+0, label = "PhiNO") +
    geom_segment(aes(x=0, y=0, xend=vectorstraitsA[3,1], yend=vectorstraitsA[3,3]), data=vectorstraitsA, 
        color="#444444", size=1.2, arrow = arrow(length = unit(0.03, "npc"))) +
        annotate("text", x = vectorstraitsA[3,1]+0.65, y = vectorstraitsA[3,3]+0.03, label = "PhiNPQ") +
    geom_segment(aes(x=0, y=0, xend=vectorstraitsA[4,1], yend=vectorstraitsA[4,3]), data=vectorstraitsA, 
        color="#444444", size=1.2, arrow = arrow(length = unit(0.03, "npc"))) +
        annotate("text", x = vectorstraitsA[4,1]-0.3, y = vectorstraitsA[4,3]-0.15, label = "ECSt") +
    geom_segment(aes(x=0, y=0, xend=vectorstraitsA[5,1], yend=vectorstraitsA[5,3]), data=vectorstraitsA, 
        color="#444444", size=1.2, arrow = arrow(length = unit(0.03, "npc"))) +
        annotate("text", x = vectorstraitsA[5,1]-0.35, y = vectorstraitsA[5,3]+0.15, label = "vH+")
    # geom_segment(aes(x=0, y=0, xend=vectorstraitsA[1,1], yend=vectorstraitsA[1,3]), data=vectorstraitsA, 
    #     color="#444444", size=1.2, arrow = arrow(length = unit(0.03, "npc"))) +
    #     annotate("text", x = gHVecA$x2-0.3, y = gHVecA$y2-0.05, label = "gH+")

traitsPlotVecA2

```

### Norm Phys PCA
```{r, echo = FALSE, results = 'hide', fig.keep = 'none'}
traitsPlotN <- 
    ggplot(PCAtraits3N, aes(x=PC1, y=PC2)) +
        coord_fixed() +  #ratio=5
        geom_point(size=3.5, aes(color=microTrt)) +  #alsen squares, norm circles?
        xlab(percentagetraitsN[1]) + ylab(percentagetraitsN[2]) +
        scale_shape_discrete("Variety") +
        scale_color_manual("",  values = micro_color) +
        theme(legend.position = "bottom", legend.margin=margin(0,0,0,0), 
              legend.box.margin=margin(-10,-10,-6,10), legend.key.width = unit(.2, "cm"),
              axis.title = element_text(size = 13), axis.text = element_text(size = 13),
              legend.text=element_text(size=10), legend.title=element_blank()) +
        guides(color = 'none') +
        #guides(color=guide_legend(nrow=2, byrow=TRUE)) +
        annotate(geom = "text", x = -2.8, y = 2.3, label = "B)\nNorm", size = 5, hjust = 0)

traitsPlotVecN <- traitsPlotN + 
    geom_segment(aes(x=0, y=0, xend=vectorstraitsN[1,1], yend=vectorstraitsN[1,2]), data=vectorstraitsN, 
        color="#444444", size=1.2, arrow = arrow(length = unit(0.03, "npc"))) +
        annotate("text", x = vectorstraitsN[1,1]-0.4, y = vectorstraitsN[1,2]-0, label = "Phi2") +
    geom_segment(aes(x=0, y=0, xend=vectorstraitsN[2,1], yend=vectorstraitsN[2,2]), data=vectorstraitsN, 
        color="#444444", size=1.2, arrow = arrow(length = unit(0.03, "npc"))) +
        annotate("text", x = vectorstraitsN[2,1]-0.22, y = vectorstraitsN[2,2]-0.13, label = "PhiNO") +
    geom_segment(aes(x=0, y=0, xend=vectorstraitsN[3,1], yend=vectorstraitsN[3,2]), data=vectorstraitsN, 
        color="#444444", size=1.2, arrow = arrow(length = unit(0.03, "npc"))) +
        annotate("text", x = vectorstraitsN[3,1]+0.50, y = vectorstraitsN[3,2]+0.20, label = "PhiNPQ") +
    geom_segment(aes(x=0, y=0, xend=vectorstraitsN[4,1], yend=vectorstraitsN[4,2]), data=vectorstraitsN, 
        color="#444444", size=1.2, arrow = arrow(length = unit(0.03, "npc"))) +
        annotate("text", x = vectorstraitsN[4,1]-0.1, y = vectorstraitsN[4,2]-0.15, label = "ECSt") +
    geom_segment(aes(x=0, y=0, xend=vectorstraitsN[5,1], yend=vectorstraitsN[5,2]), data=vectorstraitsN, 
        color="#444444", size=1.2, arrow = arrow(length = unit(0.03, "npc"))) +
        annotate("text", x = vectorstraitsN[5,1]-0.35, y = vectorstraitsN[5,2]-0.05, label = "vH+")
    # geom_segment(aes(x=0, y=0, xend=vectorstraitsN[1,1], yend=vectorstraitsN[1,2]), data=vectorstraitsN, 
    #     color="#444444", size=1.2, arrow = arrow(length = unit(0.03, "npc"))) +
    #     annotate("text", x = gHVecN$x2-0.38, y = gHVecN$y2-0.05, label = "gH+")

traitsPlotVecN
```

```{r, echo = FALSE, results = 'hide', fig.keep = 'none'}
traitsPlotN2 <- 
    ggplot(PCAtraits3N, aes(x=PC1, y=PC3)) +
        coord_fixed() +  #ratio=5
        geom_point(size=3.5, aes(color=microTrt)) +  #alsen squares, norm circles?
        xlab(percentagetraitsN[1]) + ylab(percentagetraitsN[3]) +
        scale_shape_discrete("Variety") +
        scale_color_manual("",  values = micro_color) +
        theme(legend.position = "bottom", legend.margin=margin(0,0,0,0), 
              legend.box.margin=margin(-10,-10,-6,10), legend.key.width = unit(.2, "cm"),
              axis.title = element_text(size = 13), axis.text = element_text(size = 13),
              legend.text=element_text(size=10), legend.title=element_blank()) +
        guides(color = 'none') +
        #guides(color=guide_legend(nrow=2, byrow=TRUE)) +
        annotate(geom = "text", x = -2.8, y = 2.2, label = "D)\nNorm", size = 5, hjust = 0)

traitsPlotVecN2 <- traitsPlotN2 + 
    geom_segment(aes(x=0, y=0, xend=vectorstraitsN[1,1], yend=vectorstraitsN[1,3]), data=vectorstraitsN, 
        color="#444444", size=1.2, arrow = arrow(length = unit(0.03, "npc"))) +
        annotate("text", x = vectorstraitsN[1,1]-0.4, y = vectorstraitsN[1,3]+0.05, label = "Phi2") +
    geom_segment(aes(x=0, y=0, xend=vectorstraitsN[2,1], yend=vectorstraitsN[2,3]), data=vectorstraitsN, 
        color="#444444", size=1.2, arrow = arrow(length = unit(0.03, "npc"))) +
        annotate("text", x = vectorstraitsN[2,1]-0.22, y = vectorstraitsN[2,3]-0.15, label = "PhiNO") +
    geom_segment(aes(x=0, y=0, xend=vectorstraitsN[3,1], yend=vectorstraitsN[3,3]), data=vectorstraitsN, 
        color="#444444", size=1.2, arrow = arrow(length = unit(0.03, "npc"))) +
        annotate("text", x = vectorstraitsN[3,1]+0.50, y = vectorstraitsN[3,3]+0.20, label = "PhiNPQ") +
    geom_segment(aes(x=0, y=0, xend=vectorstraitsN[4,1], yend=vectorstraitsN[4,3]), data=vectorstraitsN, 
        color="#444444", size=1.2, arrow = arrow(length = unit(0.03, "npc"))) +
        annotate("text", x = vectorstraitsN[4,1], y = vectorstraitsN[4,3]+0.15, label = "ECSt") +
    geom_segment(aes(x=0, y=0, xend=vectorstraitsN[5,1], yend=vectorstraitsN[5,3]), data=vectorstraitsN, 
        color="#444444", size=1.2, arrow = arrow(length = unit(0.03, "npc"))) +
        annotate("text", x = vectorstraitsN[5,1]-0.35, y = vectorstraitsN[5,3]-0.05, label = "vH+")
    # geom_segment(aes(x=0, y=0, xend=vectorstraitsN[1,1], yend=vectorstraitsN[1,3]), data=vectorstraitsN, 
    #     color="#444444", size=1.2, arrow = arrow(length = unit(0.03, "npc"))) +
    #     annotate("text", x = gHVecN$x2-0.38, y = gHVecN$y2-0.05, label = "gH+")

traitsPlotVecN2
```

```{r}
#tiff("./figures/PCA Phys - Var Biplot.tiff", width = 7, height = 7, units="in", res=600)
ggarrange(traitsPlotVecA, traitsPlotVecN, traitsPlotVecA2, traitsPlotVecN2, widths = c(1,1))
#dev.off()

```

```{r, echo = FALSE, results = 'hide'}
# tiff("./figures/PCA Phys - Var Biplot-poster.tiff", width = 7, height = 4, units="in", res=600)
# ggarrange(traitsPlotVecA, traitsPlotVecN2, widths = c(1,1))
# dev.off()
```

# Does Host Variety x Bacteria *indirectly* predict Disease? 
* Indirect Effects Test
* Backward model selection approach based on BIC

```{r, echo = FALSE, results = 'hide'}
str(pot2)  # 48 obs
table(pot2$hostVar, pot2$microTrt)
# 2 Var * 8 micro * 3 reps=block averages

# hist(sqrt(pot2$DONmeans))
# hist(log(pot2$FgTameans))
# hist((pot2$AUDPCmeans)) #all bimodal

```

### DON

#### ALsen
```{r}
PC.DON.A <- lm(sqrt(DONmeans) ~ PC1 + PC2 + PC3 + SPADmeans + leaf_thicknessmeans, data = PCAtraits3A)
PC.back <- step(PC.DON.A, direction="backward", k = log(24),
                    scope = c(lower = ~ 1))  #PC2 and SPADmeans clearly lowest BIC
PC.DON.A <- lm(sqrt(DONmeans) ~ PC2 + SPADmeans, data = PCAtraits3A)
Anova(PC.DON.A, type = 2)
summary(PC.DON.A)$adj.r.squared
coef(PC.DON.A)
```

#### Norm
```{r}
PC.DON.N <- lm(sqrt(DONmeans) ~ PC1 + PC2 + PC3 + SPADmeans + leaf_thicknessmeans, data = PCAtraits3N)
PC.back <- step(PC.DON.N, direction="backward", k = log(24),
                    scope = c(lower = ~ 1))  #leaf_thicknessmeans lowest BIC
PC.DON.N <- lm(sqrt(DONmeans) ~ PC3 + leaf_thicknessmeans, data = PCAtraits3N)
Anova(PC.DON.N, type = 2)
summary(PC.DON.N)$adj.r.squared
coef(PC.DON.N)
```


```{r, results = 'hide', echo = FALSE, fig.width = 8}
pcA1 <- ggplot(data = PCAtraits3A, aes(y=DONmeans, x = PC2, colour = microTrt)) +
    geom_point(size = 3) +
    scale_x_continuous("PC-2") +
    scale_y_continuous("DON", trans = "sqrt", limits = c(0,215)) +
    scale_colour_manual(values = micro_color) +
    geom_smooth(method = lm, formula = 'y ~ x', se = TRUE, col = "gray20", lty = 1) +
    annotate("text", x = 1.4, y = 200, label = "A)\nAlsen", size = 4.5, hjust = 0) +
    guides(color = 'none')
pcA2 <- ggplot(data = PCAtraits3A, aes(y=DONmeans, x = SPADmeans, colour = microTrt)) +
    geom_point(size = 3) +
    scale_x_continuous("Rel. Chlorophyll Content") +
    scale_y_continuous("DON", trans = "sqrt", limits = c(0,215)) +
    scale_colour_manual(values = micro_color) +
    geom_smooth(method = lm, formula = 'y ~ x', se = TRUE, col = "gray20", lty = 1) +
    annotate("text", x = 62.5, y = 200, label = "B)\nAlsen", size = 4.5, hjust = 0) +
    guides(color = 'none')

pcN1 <- ggplot(data = PCAtraits3N, aes(y=DONmeans, x = PC3, colour = microTrt)) +
    geom_point(size = 3) +
    scale_x_continuous("PC-3") +
    scale_y_continuous("DON", trans = "sqrt", limits = c(0,215)) +
    scale_colour_manual(values = micro_color) +
    geom_smooth(method = lm, formula = 'y ~ x', se = TRUE, col = "gray20", lty = 1) +
    annotate("text", x = 1.3, y = 200, label = "C)\nNorm", size = 4.5, hjust = 0) +
    guides(color = 'none')

# tiff("./figures/DON ~ Phys.tiff", width = 7, height = 3.2, units="in", res=600)
# multiplot(pcA1, pcA2, pcN1, cols = 3)
# dev.off()

```


### FgTa

#### ALsen
```{r}
PC.FgTa.A <- lm(log(FgTameans) ~ PC1 + PC2 + PC3 + SPADmeans + leaf_thicknessmeans, data = PCAtraits3A)
PC.back <- step(PC.FgTa.A, direction="backward", k = log(24),
                    scope = c(lower = ~ 1))  #PC2 and SPADmeans
PC.FgTa.A <- lm(log(FgTameans) ~ PC2 + SPADmeans, data = PCAtraits3A)
Anova(PC.FgTa.A, type = 2)
summary(PC.FgTa.A)$adj.r.squared
coef(PC.FgTa.A)
```

#### Norm
```{r}
PC.FgTa.N <- lm(log(FgTameans) ~ PC1 + PC2 + PC3 + SPADmeans + leaf_thicknessmeans, data = PCAtraits3N)
PC.back <- step(PC.FgTa.N, direction="backward", k = log(24),
                    scope = c(lower = ~ 1))  #model w/ PC3 & leafthickness, stop reducing AIC significantly,
# BUT! nothing really a good predictor here (significance wise, very skeptically
#   marginally significant leaf_thickness if the only thing in the model)
PC.FgTa.N <- lm(log(FgTameans) ~ PC3 + leaf_thicknessmeans, data = PCAtraits3N)
Anova(PC.FgTa.N, type = 2)
summary(PC.FgTa.N)$adj.r.squared
coef(PC.FgTa.N)
```


```{r, results = 'hide', echo = FALSE, fig.width = 8}
f.pcA1 <- ggplot(data = PCAtraits3A, aes(y=FgTameans, x = PC2, colour = microTrt)) +
    geom_point(size = 3) +
    scale_x_continuous("PC-2") +
    scale_y_continuous(expression(paste(italic("Fusarium")," Load")),
                       trans = "log", breaks = c(.2,.5,1,2,5), limits = c(.1,13)) +
    scale_colour_manual(values = micro_color) +
    geom_smooth(method = lm, formula = 'y ~ x', se = TRUE, col = "gray20", lty = 1) +
    annotate("text", x = 1.4, y = 11, label = "D)\nAlsen", size = 4.5, hjust = 0) +
    guides(color = 'none')
f.pcA2 <- ggplot(data = PCAtraits3A, aes(y=FgTameans, x = SPADmeans, colour = microTrt)) +
    geom_point(size = 3) +
    scale_x_continuous("Rel. Chlorophyll Content") +
    scale_y_continuous(expression(paste(italic("Fusarium")," Load")),
                       trans = "log", breaks = c(.2,.5,1,2,5), limits = c(.1,13)) +
    scale_colour_manual(values = micro_color) +
    geom_smooth(method = lm, formula = 'y ~ x', se = TRUE, col = "gray20", lty = 1) +
    annotate("text", x = 62.5, y = 11, label = "E)\nAlsen", size = 4.5, hjust = 0) +
    guides(color = 'none')

f.pcN1 <- ggplot(data = PCAtraits3N, aes(y=FgTameans, x = leaf_thicknessmeans, colour = microTrt)) +
    geom_point(size = 3) +
    scale_x_continuous(expression(paste("Leaf Thickness ",mu,"M"))) +
    scale_y_continuous(expression(paste(italic("Fusarium")," Load")),
                       trans = "log", breaks = c(.2,.5,1,2,5), limits = c(.1,13)) +
    scale_colour_manual(values = micro_color) +
    geom_smooth(method = lm, formula = 'y ~ x', se = TRUE, col = "gray20", lty = 1) +
    annotate("text", x = 0.50, y = 11, label = "F)\nNorm", size = 4.5, hjust = 0) +
    guides(color = 'none')
# tiff("./figures/FgTa ~ Phys biplot.tiff", width = 7, height = 4.3, units="in", res=600)
# multiplot(f.pcA1, f.pcA2,f.pcN1, cols = 3)
# dev.off()

```

```{r, results = 'hide', echo = FALSE}
#tiff("./figures/Disease ~ Phys all plots.tiff", width = 7, height = 7, units="in", res=600)
ggarrange(pcA1, pcA2, pcN1, f.pcA1, f.pcA2, f.pcN1, ncol = 3, nrow = 2)
#dev.off()
```



# End - Session Info
```{r}
sessionInfo()
```

